TAPIR: Threaded Application Programming on Intermittent Resources

### 2017-05-12 paper.tex
Added authors
Added results

%\author{

@@

%\alignauthor Amanda Marano \\
%        \affaddr{Department of Electrical and Computer Engineering}\\
%        \affaddr{Carnegie Mellon University}\\
%       \email{amarano@andrew.cmu.edu}
%\alignauthor Emily Ruppel \\
%        \affaddr{Department of Electrical and Computer Engineering}\\
%        \affaddr{Carnegie Mellon University}\\
%       \email{eruppel@andrew.cmu.edu}
%\alignauthor Neil Ryan \\
%        \affaddr{Department of Electrical and Computer Engineering}\\
%        \affaddr{Carnegie Mellon University}\\
%       \email{nryan@andrew.cmu.edu}
%}

@@

\subsection{Applications}
\subsubsection{Blinker (B)}
This application was built on top of the app-blinker-chain
example~\cite{blinker} and modified to remove the delays and LED activations, so
that it just passes data through three tasks until a maximum count is reached,
and then the thread terminates. This app was primarily used to demonstrate the
initial functionality of the system, and then to test mutexes. In both cases a
combination of \texttt{LOG} statements and final output values were used to
confirm the correct operation of the system.
\subsubsection{Cuckoo-RSA (CRSA)}
This application is a combination of a cuckoo filter that inserts and queries 64
values a Bloom filter-like data structure~\cite{Chain}, and RSA encryption of message
using 128-bit keys~\cite{Chain}. Unfortunately this is the only application
combination where merging the original applications did not produce the same
final result as running the applications separately. Several bytes at the end of
a crucial array for RSA were always overwritten when the applications were
merged, and we were unable to fix the problem. Unit tests running the RSA
application in single threaded mode through our scheduler and using our
multi-threading library produce the correct result, but any combination of RSA
and cuckoo filtering produces the incorrect result. However, RSA does run to
completion, so we believe the results from C-RSA are still useful.
\subsubsection{Cuckoo-Bitcount (CB)}
This application is a combination of the cuckoo filter described above and a
application that uses different algorithms to count the number of bits in an
array of integers~\cite{bitcount}. CB is a correctly working combination of two
applications that produces the same output as separate executions of the
unmodified versions of these applications. It expands the reliability of our
multi-threading library beyond simple applications like Blinker.
\subsubsection{Cuckoo-Bitcount-Blinker(CBB)}
CBB combines the cuckoo filter and bit count applications with two threads of
the blinker application. This results in four active threads at once, which is
the maximum number of threads our prototype implementation allows. We confirmed
that all four threads are fairly serviced by the round robin scheduler and they
correctly start and end.

Memory overhead (read off compilation)\\
Performance (\# reboots)\\

>>>>

@@

\author{
\alignauthor Amanda Marano \\
        \affaddr{Department of Electrical and Computer Engineering}\\
        \affaddr{Carnegie Mellon University}\\
       \email{amarano@andrew.cmu.edu}
\alignauthor Emily Ruppel \\
        \affaddr{Department of Electrical and Computer Engineering}\\
        \affaddr{Carnegie Mellon University}\\
       \email{eruppel@andrew.cmu.edu}
\alignauthor Neil Ryan \\
        \affaddr{Department of Electrical and Computer Engineering}\\
        \affaddr{Carnegie Mellon University}\\
       \email{nryan@andrew.cmu.edu}
}

@@

\subsection{Applications}
\subsubsection{Blinker (B)}
This application was built on top of the app-blinker-chain
example~\cite{blinker} and modified to remove the delays and LED activations, so
that it just passes data through three tasks until a maximum count is reached,
and then the thread terminates. This app was used to demonstrate the
initial functionality of the system, and then to test mutexes. In both cases a
combination of \texttt{LOG} statements and final output values were used to
confirm the correct operation of the system, however we do not report memory
usage or reboots during execution for this application. 
\subsubsection{Cuckoo-RSA (CRSA)}
This application is a combination of a cuckoo filter that inserts and queries 64
values a Bloom filter-like data structure~\cite{Chain}, and RSA encryption of message
using 128-bit keys~\cite{Chain}. Unfortunately this is the only application
combination where merging the original applications did not produce the same
final result as running the applications separately. Several bytes at the end of
a crucial array for RSA were always overwritten when the applications were
merged, and we were unable to fix the problem. Unit tests running the RSA
application in single threaded mode through our scheduler and using our
multi-threading library produce the correct result, but any combination of RSA
and cuckoo filtering produces the incorrect result. However, RSA does run to
completion, so we believe at least the memory overhead results from C-RSA are still useful.
\subsubsection{Cuckoo-Bitcount (CB)}
This application is a combination of the cuckoo filter described above and a
application that uses different algorithms to count the number of bits in an
array of integers~\cite{bitcount}. CB is a correctly working combination of two
applications that produces the same output as separate executions of the
unmodified versions of these applications. It expands the reliability of our
multi-threading library beyond simple applications like Blinker.
\subsubsection{Cuckoo-Bitcount-Blinker(CBB)}
CBB combines the cuckoo filter and bit count applications with two threads of
the blinker application. This results in four active threads at once, which is
the maximum number of threads our prototype implementation allows. We confirmed
that all four threads are fairly serviced by the round robin scheduler and they
correctly start and end.

\subsection{Results}
CRSA, CB, and CBB were ported from multithreaded applications to serial
applications. Rather than using the scheduler to switch between tasks, the an
application transitions to the first task of the successive application upon
completion. The memory overhead of RSA and Bitcount when they are instrumented
to run as a single thread with the additional overhead of the scheduler is shown
to provide a reference for the baseline cost of the scheduler.
Figure~\ref{fig:mem} shows the increase in code size with the switcht to the
multithreading library. Approximately 3 lines or code in user space needed to be
added to each of the multithreaded applications, but the library code is
significant. The fraction of overhead ranges from 12.6\% for CB to nearly 30\%
for bitcount.

\begin{figure}[h]
  \centering
  \includegraphics[width=0.99\columnwidth]{figs/memory}
  \caption{{\bf Increase in code size}}
  \label{fig:copy}
\end{figure}

We measure the performance of the application in terms of the number of reboots
that occur before the application completes. This is an accurate measurement of
performanc in intermittent systems because the majority of the execution time is
spent recharging the capacitors. In our system set up, as shown in Figure
~\ref{fig:setup}, the capacitor charging time is consistent at approximately 3.5
seconds. We only report reboot data for CBB and CB because of the discrepancy in
CRSA output between the multithreaded and serial versions.
Figure~\ref{fig:reboots} shows that the scheduler has a significant cost to the
system. The multithreaded executions incur significantly more reboots than the
serial versions. This is to be expected due to the additional instructions of
scheduler code and additional writes to non-volatile memory required to access
the scheduler. 
\begin{figure}[h]
  \centering
  \includegraphics[width=0.99\columnwidth]{figs/setup}
  \caption{{\bf Test setup using a Capybara board, and RF power source}}
  \label{fig:setup}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[width=0.99\columnwidth]{figs/reboots}
  \caption{{\bf Number of reboots per execution}}
  \label{fig:reboots}
\end{figure}
